\vspace{1.0cm} \LARGE \bf MERS-CoV spillover at the camel-human interface

### 2017-11-21 mers-structure.tex
Address #29

Overall, our analyses indicate that MERS-CoV is poorly suited for human-to-human transmission, with an estimated $R_{0}<0.90$ and sequence sampling likely to be biased towards large hospital outbreaks (Figure \ref{mers_epi_grid}).
All matching simulations exhibit highly skewed distributions of case cluster sizes with long tails (Figure \ref{cluster_distributions}), which is qualitatively similar to the results of \citep{cauchemez_unraveling_2016}.
Given these findings, and the fact that large outbreaks of MERS occurred in hospitals, the combination of frequent spillover of MERS-CoV into humans and occasional outbreak amplification via poor hygiene practices \citep{assiri_hospital_2013,chen_comparative_2017} appear sufficient to explain observed epidemiological patterns of MERS-CoV.

@@

\begin{figure}[h]
\centering
	\includegraphics[width=0.75\textwidth]{figures/mers_epi_grid.png}
	\caption{\textbf{Monte Carlo simulations of human transmission clusters.}
From top to bottom each row corresponds to departures from completely random sequencing efforts with respect to case cluster size (bias parameter=1.0) to sequencing increasingly biased towards capturing large case clusters (bias=2.0, bias=3.0).
Leftmost scatter plots show the distribution of individual Monte Carlo simulation sequence cluster size statistics (mean and skewness) coloured by the $R_{0}$ value used for the simulation.
The dotted rectangle identifies the 95\% highest posterior density bounds for sequence cluster size mean and skewness observed for empirical MERS-CoV data.
The distribution of $R_{0}$ values matching empirical data are shown in the middle, on the same $y$-axis across all levels of the bias parameter.
Under unbiased sequencing (bias=1.0) only 0.45\% of simulations fit our phylogenetic observations, while 1.79\% and 1.67\% of simulations fit for bias levels of 2.0 and 3.0, respectively.
Correspondingly, we estimate 11.6\% support for a model with bias level 1.0, 45.7\% support for a model with bias level 2.0, and 42.7\% support for a model with bias level 3.0.
Bins falling inside the 95\% percentiles are coloured by $R_{0}$, as in the leftmost scatter plot.
While the 95\% percentiles for $R_{0}$ values are close to 1.0 (0.71--0.98) for the unbiased sequencing simulation (\textit{i.e.}\ uniform sequencing efforts, in which every case is equally likely to be sequenced), we also note that increasing levels of bias are considerably more to likely to generate MERS-CoV-like sequence clusters.

The distribution of total number of introductions associated with simulations matching MERS-CoV sequence clusters is shown in the plots on the right, on the same $y$-axis across all levels of bias.
Darker shade of grey indicates bins falling within the 95\% percentiles.
The median number of cross-species introductions observed in simulations matching empirical data without bias are 346 (95\% percentiles 262--439).
These numbers jump up to 568 (95\% percentiles 430--727) for bias = 2.0 and 656 (95\% percentiles 488--853) for bias = 3.0 simulations.
Model averaging would suggest plausible numbers of introductions between 311 and 811.
	}
	\label{mers_epi_grid}
\end{figure}

>>>>

Overall, our analyses indicate that MERS-CoV is poorly suited for human-to-human transmission, with an estimated $R_{0}<0.90$ and sequence sampling likely to be biased towards large hospital outbreaks (Figure \ref{mers_epi_grid}).
All matching simulations exhibit highly skewed distributions of case cluster sizes with long tails (Figure \ref{cluster_distributions}), which is qualitatively similar to the results of \citep{cauchemez_unraveling_2016}.
We find that simulated sequence cluster sizes resemble observed sequence cluster sizes in the posterior distribution, with a slight reduction in mid-sized clusters in simulated data (Figure \ref{qq}).
Given these findings, and the fact that large outbreaks of MERS occurred in hospitals, the combination of frequent spillover of MERS-CoV into humans and occasional outbreak amplification via poor hygiene practices \citep{assiri_hospital_2013,chen_comparative_2017} appear sufficient to explain observed epidemiological patterns of MERS-CoV.

@@

\begin{figure}[h]
\centering
	\includegraphics[width=0.65\textwidth]{figures/mers_qq.png}
	\caption{\textbf{Quantile-quantile (Q-Q) plot of empirical and simulated sequence cluster sizes.}
Density of sequence cluster size percentiles (1st--99th, calculated across a grid of 50 values) calculated for random states from the posterior distribution ($x$-axis) and matching simulations ($y$-axis).
Most values fall on the one-to-one line, with a heavier tail in mid-sized sequence clusters in empirical data, manifesting as a greater density of points below the one-to-one line in the middle.
	}
	\label{qq}
\end{figure}

\begin{figure}[h]
\centering
	\includegraphics[width=0.75\textwidth]{figures/mers_epi_grid.png}
	\caption{\textbf{Monte Carlo simulations of human transmission clusters.}
From top to bottom each row corresponds to departures from completely random sequencing efforts with respect to case cluster size (bias parameter=1.0) to sequencing increasingly biased towards capturing large case clusters (bias=2.0, bias=3.0).
Leftmost scatter plots show the distribution of individual Monte Carlo simulation sequence cluster size statistics (mean and skewness) coloured by the $R_{0}$ value used for the simulation.
The dotted rectangle identifies the 95\% highest posterior density bounds for sequence cluster size mean and skewness observed for empirical MERS-CoV data.
The distribution of $R_{0}$ values matching empirical data are shown in the middle, on the same $y$-axis across all levels of the bias parameter.
Under unbiased sequencing (bias=1.0) only 0.45\% of simulations fit our phylogenetic observations, while 1.79\% and 1.67\% of simulations fit for bias levels of 2.0 and 3.0, respectively.
Correspondingly, we estimate 11.6\% support for a model with bias level 1.0, 45.7\% support for a model with bias level 2.0, and 42.7\% support for a model with bias level 3.0.
Bins falling inside the 95\% percentiles are coloured by $R_{0}$, as in the leftmost scatter plot.
While the 95\% percentiles for $R_{0}$ values are close to 1.0 (0.71--0.98) for the unbiased sequencing simulation (\textit{i.e.}\ uniform sequencing efforts, in which every case is equally likely to be sequenced), we also note that increasing levels of bias are considerably more to likely to generate MERS-CoV-like sequence clusters.
The distribution of total number of introductions associated with simulations matching MERS-CoV sequence clusters is shown in the plots on the right, on the same $y$-axis across all levels of bias.
Darker shade of grey indicates bins falling within the 95\% percentiles.
The median number of cross-species introductions observed in simulations matching empirical data without bias are 346 (95\% percentiles 262--439).
These numbers jump up to 568 (95\% percentiles 430--727) for bias = 2.0 and 656 (95\% percentiles 488--853) for bias = 3.0 simulations.
Model averaging would suggest plausible numbers of introductions between 311 and 811.
  }
	\label{mers_epi_grid}
\end{figure}
